Pool Model - A unified Mathematical\\ Modelling Framework for Food Science

### 2025-09-03 paper/pool_model.tex
write abstract

\begin{abstract}\noindent
    \textbf{Notes from Christian}
    Different empirical and inactivation models applied in the estimation
    of parameters and modelling of phenomena in food microbiology, gen-
    eralized Lotka Voltera approach used to describe the interaction of mi-
    croorganisms, mechanistic models utilized to provide more insight into
    the underlying dynamic can all be unified as variations of a new model
    known as three pool model. Compared against the aforementioned
    methods which are applicable in specific applications with some limi-
    tations, the present model turns out to be less complex with general
    applicability, low number of parameters and preferably to explain the
    mechanism of the phenomenon of interest. To illustrate the generality
    of the model, three different examples are reported.
\end{abstract}

>>>>

\begin{abstract}\noindent
    Bacterial consortia are complex systems which are often comprised of multiple distinct species.
    Their dynamics on a population-based level often result from interactions that lead to
    inter-species cooperation or competition.
    Various empirical and inactivation models have been conceived and applied to problems related to
    food microbiology.
    However, the utilized models live on the popluation level, omitting any spatial effects such as
    heterogeneity or diffusion.
    In this work we present a novel generalized framework, called "pool model" which uses multiple
    distinct pools that can hold bacterial states of distinct species or extracellular nutrient
    abundance.
    It allows us to describe these systems mechanistically as variations of an underlying, more
    fundamental model.
    To illustrate the usefulness of the model, multiple different examples are reported.
    Finally, we construct an \acl{abm} for a given pool model, showcasing that the resulting model
    arises naturally from the individual behaviour of many cellular agents.
    We further apply this \ac{abm} to study the role of spatial effects on the dynamcis of bacterial
    growth.
    % \textbf{Notes from Christian}
    % Different empirical and inactivation models applied in the estimation
    % of parameters and modelling of phenomena in food microbiology, gen-
    % eralized Lotka Voltera approach used to describe the interaction of mi-
    % croorganisms, mechanistic models utilized to provide more insight into
    % the underlying dynamic can all be unified as variations of a new model
    % known as three pool model. Compared against the aforementioned
    % methods which are applicable in specific applications with some limi-
    % tations, the present model turns out to be less complex with general
    % applicability, low number of parameters and preferably to explain the
    % mechanism of the phenomenon of interest. To illustrate the generality
    % of the model, three different examples are reported.
\end{abstract}

\noindent
\textbf{Keywords:}
Bacterial Consortia, Competition and Cooperation, Mechanistic Modeling, Spatial Effects
